---
title: "Raman spectra data analysis"
author: "Andres Beltran"
date: '2022-07-09'
output: rmdformats::downcute
---


# Loading the data

As received the spectra is separated in different folders:

```{r}
dir('./degradation of rods_Yohanna/')
```

Two folders for two samples,  inside each folder we have:

* First, for `309_1`:

```{r}
dir('./degradation of rods_Yohanna/309_1/')
```
* Second, for `309_2`:

```{r}
dir('./degradation of rods_Yohanna/309_4/')
```

Now we know that there are two replicates at 50 mW and two at 5 mW for sample `309_1`, and three replicates for `309_2` without further information depicted in the folder names. Nonetheless, each file name specifieas that all spectra for 309_4 is taken using 5 mW.

The names that lack extension are folders, so within each folder we will find the file of the spectrum:

```{r}
dir('./degradation of rods_Yohanna/309_1/spectrum 1, 50mW/')
```
```{r}
files <- list.files('./degradation of rods_Yohanna/', pattern = 'Data',recursive = TRUE)
files
```

Once we know where the data files are, we can import them as follows:

```{r}
files <- paste0('./degradation of rods_Yohanna/',files)
data <- lapply(files, read.table, header = F)
```


Once we have all the data we need in the object `data` we can transform it to examine it.

First we check if the x axis is the same for all spectra:

1 and 2: 

```{r}
sum(data[[1]][,1] == data[[2]][,1])
nrow(data[[1]])
```

```{r}

matches <- vector('numeric', length = 7)
for (i in 1:7 ) {
  
 matches[i] <- sum(data[[1]][,1] == data[[i]][,1]) 
  
  
}
matches
```

Once we have verified that the x axis is the same for all of the spectra, we can extract it and save it separatelly as follows:


```{r}
shift <- as.numeric(data[[1]][,1])
```

and then, create a data.frame with all of the y axes for all samples:

```{r}
df1 <- as.data.frame(lapply(data, '[', 2))
df1 <- as.data.frame(t(df1))
colnames(df1) <- shift

```

to assign a shorter name for each sample, we can inspect and extract useful information from the file names `files`:


```{r}
files
```

```{r}
names <- c('309_1_1 50 mW',
           '309_1_2 50 mW',
           '309_1_3 5 mW',
           '309_1_4 5 mW',
           '309_4_1 5 mW',
           '309_4_2 5 mW',
           '309_4_3 5 mW'
           )
```

We now assing the sample names to the rows:

```{r}
rownames(df1) <- names
```

Since we received the x axis in nm, we proceed to express it in 1/cm:





$$\bar{\nu} = \frac{1}{\lambda}\\$$

$$\bar{\nu} \;[1/cm] = \frac{1}{\lambda [nm]} \\$$

$$cm = 1 nm \: \cdot \: (\frac{1 m}{10^9 nm}) \: \cdot \: (\frac{10^2 cm}{1 m})\\$$

$$cm^{-1} = (1 nm \: \cdot \: (\frac{1 m}{10^9 nm}) \: \cdot \: (\frac{10^2 cm}{1 m}))^{-1} \\$$

$$cm^{-1} = (\frac{1}{10^7})^{-1} \: \cdot \:  nm \\$$

$$cm^{-1} = 10^7 \: \cdot \:  nm$$





```{r}
shift2 <- 1/(shift/10^7)
```

The `range`s  of the x axes do not match with the expected values:

```{r}
range(shift)# allegedly nm
range(shift2) # 1/cm from nm
```

So we create a new axis for sake of plotting:

```{r}
shift3 <- seq(200,4000, length = 1024)
```


For the plot:

```{r}

col <- c(rep('red',2),rep('pink',2),rep('black',3))

for(i in 1:7){
  plot(shift3,
       df1[i,],
       type = 'l',
       col = col[i],
       xlim =,
       ylim =,
       xlab = '',
       ylab = '',
       axes = FALSE
  
  )
    par(new =T)
}
axis(1)
axis(2)
box()

legend('topleft',names, col = col, lty = 1, cex= 0.5)

```
At this moment we can see that replicates are very different between each other.


## Baseline correction

Since all spectra are very different between each other (even between replicates, a test can be found [here](https://github.com/afbeltranr/RamanCLiMo/blob/main/test/test.R)) it will be more useful if the estimation of the baseline is applied to each spectrum separately. 

To do this iteratively, we can create a list to save each data frame in one slot of the list `listOfDf`:

```{r}
listOfDf <- vector('list', 7)
```


Now we create a data.frame in which we will save two parameters of baseline correction: 

* One is the `noise` threshold at which the function recognizes the noise. 
* The second one is the degrees of freedom used by the `smooth.spline` function available in the package `stats`. If not specified leave-one-out cross-validation (ordinary or ‘generalized’ as determined by cv) is used to determine smoothing parameters. 

These parameters have been selected subjectively seeking for a better appearance of the spectra. If further analysis and information extraction were to be performed to this data set, these parameters should be selected via optimization using the correct objective function as optimization criteria.




```{r, warnings = F}
library(hyperSpec)
param <- data.frame(noise = c(10,17,8,8,200,15,50),
                      df. = c(15,17,10,10,5,20,5))

for(i in 1:7){

listOfDf[[i]] <- df1[i,]
  
spc[[i]] <-            new('hyperSpec',
                       spc = listOfDf[[i]][, 50:1023],
                       wavelength = shift3[50:1023]) 

bend[[i]]   <-         wl.eval(spc[[i]],
                        function (x)
                        x^6+x^5+x^4+x^3+x^2,
                        normalize.wl =normalize01)

 bl[[i]]  <-        spc.rubberband(spc[[i]]+bend[[i]],
                     noise =param$noise[i],
                     df = param$df.[i]
                     )-bend[[i]]
 
 spc2[[i]] <- spc[[i]] - bl[[i]]
 
   plot(spc2[[i]], col = T)
 #  par(new  = T)

plot(spc[[i]],add=TRUE, wl.reverse = F)
plot(bl[[i]], add=TRUE, col=2,wl.reverse =F )
  
}

a <- matrix(nrow  = 7, ncol = 974)
 
 
for(i in  1:7){
  
 a[i,] <- t(as.data.frame(spc2[[i]][,1]$spc[1,]))
  

 
}

 colnames(a) <- as.numeric(colnames(as.data.frame(spc2[[1]][,1]$spc)))
```

```{r}
 for(i in c(1:7)[-5]){
   
 plot(colnames(a),
     a[i,],
     type = 'l',
     # axes = 'F',
     ylab = '',
     xlab = '',
      ylim = c(0,200),
     # xlim = c()
     col = col[i]
     
     )  
   par(new = T)
 }


```
```{r, message = F, warning = F, echo = F}
 library(hyperSpec)
library(dplyr)
library(prospectr)


w2 <-70
sg1 <- sg <- matrix(ncol = ncol(a)-w2, nrow= nrow(a))
sg1 <- as.data.frame(sg1) 

sg1 <- savitzkyGolay(X = a
                        ,m = 2  ,
                        p = 7,
                     w = w2+1)
for(i in 1:7){

  plot(as.numeric(colnames(sg1)),
       sg1[i,],
       xlab = 'raman shift',
       ylab = 'smoothed spectra',
       # xlim = c(1700,790),
             ylim = c(-15,15),
       type = 'l',
       col = col[i])

  
    par(new = T)
}
```


```{r}
sessionInfo()
```


