---
title: "Raman spectra data analysis"
author: "Andres Beltran"
date: '2022-07-09'
output: rmdformats::downcute
---


# Loading the data

As received the spectra is separated in different folders:

```{r}
dir('./degradation of rods_Yohanna/')
```

Two folders for two samples,  inside each folder we have:

* First, for `309_1`:

```{r}
dir('./degradation of rods_Yohanna/309_1/')
```
* Second, for `309_2`:

```{r}
dir('./degradation of rods_Yohanna/309_4/')
```

Now we know that there are two replicates at 50 mW and two at 5 mW for sample `309_1`, and three replicates for `309_2` without further information depicted in the folder names. Nonetheless, each file name specifieas that all spectra for 309_4 is taken using 5 mW.

The names that lack extension are folders, so within each folder we will find the file of the spectrum:

```{r}
dir('./degradation of rods_Yohanna/309_1/spectrum 1, 50mW/')
```
```{r}
files <- list.files('./degradation of rods_Yohanna/', pattern = 'Data',recursive = TRUE)
files
```

Once we know where the data files are, we can import them as follows:

```{r}
files <- paste0('./degradation of rods_Yohanna/',files)
data <- lapply(files, read.table, header = F)
```


Once we have all the data we need in the object `data` we can transform it to examine it.

First we check if the x axis is the same for all spectra:

1 and 2: 

```{r}
sum(data[[1]][,1] == data[[2]][,1])
nrow(data[[1]])
```

```{r}

matches <- vector('numeric', length = 7)
for (i in 1:7 ) {
  
 matches[i] <- sum(data[[1]][,1] == data[[i]][,1]) 
  
  
}
matches
```

Once we have verified that the x axis is the same for all of the spectra, we can extract it and save it separatelly as follows:


```{r}
shift <- as.numeric(data[[1]][,1])
```

and then, create a data.frame with all of the y axes for all samples:

```{r}
df1 <- as.data.frame(lapply(data, '[', 2))
df1 <- as.data.frame(t(df1))
colnames(df1) <- shift

```

to assign a shorter name for each sample, we can inspect and extract useful information from the file names `files`:


```{r}
files
```

```{r}
names <- c('309_1_1 50 mW',
           '309_1_2 50 mW',
           '309_1_3 5 mW',
           '309_1_4 5 mW',
           '309_4_1 5 mW',
           '309_4_2 5 mW',
           '309_4_3 5 mW'
           )
```

We now assing the sample names to the rows:

```{r}
rownames(df1) <- names
```

Since we received the x axis in nm, we proceed to express it in 1/cm:





$$\bar{\nu} = \frac{1}{\lambda}\\$$

$$\bar{\nu} \;[1/cm] = \frac{1}{\lambda [nm]} \\$$

$$cm = 1 nm \: \cdot \: (\frac{1 m}{10^9 nm}) \: \cdot \: (\frac{10^2 cm}{1 m})\\$$

$$cm^{-1} = (1 nm \: \cdot \: (\frac{1 m}{10^9 nm}) \: \cdot \: (\frac{10^2 cm}{1 m}))^{-1} \\$$

$$cm^{-1} = (\frac{1}{10^7})^{-1} \: \cdot \:  nm \\$$

$$cm^{-1} = 10^7 \: \cdot \:  nm$$





```{r}
shift2 <- 1/(shift/10^7)
```

The `range`s  of the x axes do not match with the expected values:

```{r}
range(shift)# allegedly nm
range(shift2) # 1/cm from nm
```

So we create a new axis for sake of plotting:

```{r}
shift3 <- seq(200,4000, length = 1024)
```


For the plot:

```{r}

col <- c(rep('red',2),rep('pink',2),rep('black',3))

for(i in 1:7){
  plot(shift3,
       df1[i,],
       type = 'l',
       col = col[i],
       xlim =,
       ylim =,
       xlab = '',
       ylab = '',
       axes = FALSE
  
  )
    par(new =T)
}
axis(1)
axis(2)
box()

legend('topleft',names, col = col, lty = 1, cex= 0.5)

```
At this moment we can see that replicates are very different between each other.


## Baseline correction


```{r, message = F}
 library(hyperSpec)

spc <- new('hyperSpec',
           spc = df,
           wavelength = shift3
  
)
```

```{r}
bend <-  wl.eval(spc,
                      function (x) 
                      x^6+x^5+x^4+x^3+x^2,
                      normalize.wl = 
                        normalize01)
```


```{r, message = F, warning = F}
bl <- spc.rubberband(spc+bend,
                     noise = 5, 
                     df = 20)-bend
```


```{r}
plot(spc, wl.reverse = TRUE)
plot(bl, add=TRUE, col=2,wl.reverse = TRUE)
```

